### 2023-03-14 main.tex
dualityUpdate

\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}\label{edgeHamiltonian}
H_{x,y}=\sum_{k,l=1}^{N}\left(E_{kl}^{x}\otimes E_{lk}^{y}-E_{ll}^{x}\otimes E_{kk}^{y}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{k,l=1}^{N}\alpha_{k}^{x}\left(E_{kl}^{x}-E_{ll}^{x}\right)
\end{equation}
Here $E_{kl}^{x}$ is a copy of $E_{kl}$ defined in \eqref{actionE} acting on site $x$. 
Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

@@

H_{x}^{T}=\sum_{k,l=0}^{N}\alpha_{k}^{x}\left(E_{k\ell}^{x}-E_{\ell\ell}^{x}\right)^{T}=R_{x}\sum_{k,l=0}^{N}\alpha_{k}^{x}\left(E_{lk}^{x}-E_{ll}^{x}\right)R_{x}^{-1}

@@

    &\exp{(-E^{x})}\sum_{k=1}^{N}\sum_{l=1}^{N}\alpha_{k}^{x}\left(E_{lk}^{x}-E_{ll}^{x}\right)\exp{(E^{x})}\\&=
     \exp{(-E^{x})}\sum_{k=1}^{N}\alpha_{k}^{x}\left(E_{1k}^{x}-E_{11}^{x}\right)\exp{(E^{x})}
     + \exp{(-E^{x})}\sum_{k=1}^{N}\sum_{l=2}^{N}\alpha_{k}^{x}\left(E_{lk}^{x}-E_{ll}^{x}\right)\exp{(E^{x})}

@@

     \sum_{k=2}^{N}\alpha_{k}^{x}\left(E_{1k}+E_{11}^{x}-\sum_{a=2}^{N}(E_{ak}^{x}+E_{a1}^{x})-E_{11}^{x}+\sum_{j=2}^{N}E_{j1}^{x}\right)+\sum_{k=1}^{N}\sum_{l=2}^{N}\alpha_{k}^{x}\left(E_{lk}^{x}+E_{\ell 1}^{x}-E_{\ell\ell}^{x}-E_{l1}^{x}\right)

@@

The duality result can be state also in terms of infinitesimal generators \eqref{Generator}. In this case, the dual process, is denoted by $(\bm{\xi}(t))_{t\geq 0}$, where for each $t\geq 0$ and for each $x\in V$ we denote

@@

where we denote $r_{k}^{x}$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at site $x$ and with $m_{l}^{\widetilde{x}}$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at extra-site $\widetilde{x}$ connected to $x\in V$. This dual process has values in the enlarged state space

@@

The duality relation \eqref{DualityRelation} reads

@@

	 &\langle n^{x}|D_{x}\otimes D_{\widetilde{x}}|\xi^{x}\rangle\\&=\langle n_{1}^{x},\ldots,n_{N}^{x}|\sum_{m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\varepsilon(x)}}\langle m_{2},\ldots,m_{N}|_{x}

@@

on the other hand, on the site $x$ we have 

@@

where we applied the orthogonality relations \eqref{ortho} and the fact that $r_{1}^{x}=\nu-\sum_{k=2}^{N}r_{k}^{x}$. Finally, by taking the product over $x\in V$ we prove \eqref{dualityElements}. \newline \\
\textbf{Remark}: in the reverible situation, i.e. when $\forall x\in V$ $\alpha_{k}^{x}=\alpha_{k}$, the average of the element of the duality matrix with respect to the reversible measure \eqref{reversibleMeasure}, satisfies

@@

In \cite{UPHILL}, a two-species hard-core reaction diffusion process has been introduced. The feature of this process is that its action on the occupation variable of any species gives a difference-differential equation where both discrete Laplacians and a linear reaction term are present (multiplied for proper constants). In this subsection we write a generalization to $N\in \mathbb{N}$ species with arbitrary maximal occupation per site, again called $\nu\in\mathbb{N}_{0}$. The generator reads as follows

@@

    \mathcal{L}_{x,u}^{i}f(n)=\sum_{k,l=1}^{N}n_{k}^{x}n_{l}^{y}\left(f(n-\delta_{k}^{x}+\delta_{g_{i}(l)}^{x}+\delta_{g_{i}(k)}^{y}-\delta_{l}^{y})-f(n)\right)

@@

    \mathcal{L}_{x,y}^{m}f(n)=\sum_{l,k=0}^{N}n_{k}^{x}\left(f(n-\delta_{k}^{x}+\delta_{l}^{x})-f(n)\right)

@@

where with the symbol $\Delta_{N}$ we denote the discrete Laplacian operator on the chain, i.e. $\Delta_{L}n_{k}^{x}=n_{k}^{x+1}+n_{k}^{x-1}-2n_{k}^{x}$, $\forall x\in \{2,\ldots,L-1\}$ and $\forall k\in \{2,\ldots,N\}$. In the spirit of \cite{} we can write he the Master's equation for the evolution of the average density occupation, obtaining a system of $N-1$ difference differential equation where the discrete Laplacian of all species are coupled and linear reaction term are present. \\ 

@@

where the duality matrix $D$ is again \eqref{dualityMatrix}. Following subsection \ref{subsectionDualityProof}, to prove \eqref{DualityRelationRD} we only need tho show that the operator $S_{x,y}$ defined in \eqref{symmetry} commutes with $\sum_{i=1}^{N}H_{x,y}^{i}$ and with $H_{x,y}^{m}$. This result will follow from the commutator between $\sum_{i=1}^{N}H_{x,y}^{i}$ and $\sum_{a=2}^{N}(E_{a1}^{x})\otimes \sum_{b=2}^{N}E_{b,1}^{x}$ and from the commutator between $H_{x,y}^{m}$ and $\sum_{a=2}^{N}(E_{a1}^{x})\otimes \sum_{b=2}^{N}E_{b,1}^{x}$. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 

@@

Therefore, following a proof similar to the one in subsection \ref{subsectionDualityProof}, \eqref{DualityRelationRD} can be proved. 

>>>>

\begin{equation}\label{OriginalHamiltonian}
	\begin{split}
		H=\sum_{x,y\in \mathcal{E}}\omega_{x,y}H_{x,y}+\sum_{x\in V}\Gamma_{x}H_{x}
	\end{split}
\end{equation}
where the edge Hamiltonian is
\begin{equation}\label{edgeHamiltonian}
H_{x,y}=\sum_{k,\ell=1}^{N}\left(E_{k\ell}^{x}\otimes E_{\ell k}^{y}-E_{\ell\ell}^{x}\otimes E_{kk}^{y}\right)
 \end{equation}
 and where the site Hamiltonian is
 \begin{equation}\label{siteHamiltonian}
H_{x}=\sum_{k,\ell=1}^{N}\alpha_{k}^{x}\left(E_{k\ell}^{x}-E_{\ell\ell}^{x}\right)
\end{equation}
Here $E_{kl}^{x}$ is a copy of $E_{kl}$ defined in \eqref{actionE} acting on site $x$. 
Following  \cite{belitsky2015self}, the Hamiltonian and the genertor are linked by
\begin{equation}
H=\mathcal{L}^{T}
\end{equation}
The generator can also be expressed as

@@

H_{x}^{T}=\sum_{k,l=0}^{N}\alpha_{k}^{x}\left(E_{k\ell}^{x}-E_{\ell\ell}^{x}\right)^{T}=R_{x}\sum_{k,\ell=0}^{N}\alpha_{k}^{x}\left(E_{\ell k}^{x}-E_{\ell\ell}^{x}\right)R_{x}^{-1}

@@

    &\exp{(-E^{x})}\sum_{k=1}^{N}\sum_{\ell=1}^{N}\alpha_{k}^{x}\left(E_{\ell k}^{x}-E_{\ell\ell}^{x}\right)\exp{(E^{x})}\\&=
     \exp{(-E^{x})}\sum_{k=2}^{N}\alpha_{k}^{x}\left(E_{1k}^{x}-E_{11}^{x}\right)\exp{(E^{x})}
     + \exp{(-E^{x})}\sum_{k=1}^{N}\sum_{\ell=2}^{N}\alpha_{k}^{x}\left(E_{\ell k}^{x}-E_{\ell\ell}^{x}\right)\exp{(E^{x})}

@@

     \sum_{k=2}^{N}\alpha_{k}^{x}\left(E_{1k}+E_{11}^{x}-\sum_{a=2}^{N}(E_{ak}^{x}+E_{a1}^{x})-E_{11}^{x}+\sum_{a=2}^{N}E_{a1}^{x}\right)+\sum_{k=1}^{N}\sum_{\ell =2}^{N}\alpha_{k}^{x}\left(E_{\ell k}^{x}+E_{\ell 1}^{x}-E_{\ell\ell}^{x}-E_{\ell 1}^{x}\right)

@@

The duality result can be also stated  in terms of the infinitesimal generators \eqref{Generator}. The dual process is $(\bm{\xi}(t))_{t\geq 0}$ . For each $x\in V$ we define the $x-th$ component as

@@

 where we denote $r_{k}^{x}$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at site $x$ and with $m_{l}^{\widetilde{x}}$ the occupation of dual particle of type $k\in \{1,\ldots,N\}$ at extra-site $\widetilde{x}$ connected to $x\in V$. This dual process lives in the enlarged state space

@@

The duality relation \eqref{DualityRelation} reads as

@@

	 &\langle n^{x}|D_{x}\otimes D_{\widetilde{x}}|\xi^{x}\rangle\\&=\langle n_{1}^{x},\ldots,n_{N}^{x}|\sum_{m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}=0}^{\infty}\prod_{k=2}^{N}\left(\frac{\alpha_{k}}{\sum_{i=1}^{N}\alpha_{i}^{x}}\right)^{m_{k}^{\widetilde{x}}}\langle m_{2}^{\widetilde{x}},\ldots,m_{N}^{\widetilde{x}}|

@@

On the other hand, on the site $x$, we have 

@@

where we applied the orthogonality relations \eqref{ortho} and the fact that $r_{1}^{x}=\nu-\sum_{k=2}^{N}r_{k}^{x}$. Finally, by taking the product over $x\in V$ we prove \eqref{dualityElements}.\\
\textbf{Remark}: It is important to notice that the denomiator the elements of the duality matrix \eqref{dualityElements} $(n_{k}^{x}-r_{k}^{x})!$,  implies that, at each site $x\in V$ and for every species $k\in \{2,\ldots,N\}$ the number of dual particles is lower or equal than the number of original particles. This means that the dynamics of the dual proces is simpler than the one of the original one. It is possible to perform similar computations choosing as a symmetry $E^{x}=\sum_{a=2}^{N}E_{1a}^{x}$ or $E^{x}=\sum_{a=2}^{N}E_{aa}^{x}$. However, these would lead to duality matrices where the number of dual particles is greater than the number of original ones loosing the simplification of the dynamics. 
\textbf{Remark}: in the reverible situation, i.e. when $\forall x\in V$ $\alpha_{k}^{x}=\alpha_{k}$, the average of the element of the duality matrix with respect to the reversible measure \eqref{reversibleMeasure}, is

@@

In \cite{casini2022uphill}, a two-species hard-core reaction diffusion process has been introduced. The feature of this process is that its action on the occupation variable of any species gives a difference-differential equation where both discrete Laplacians and a linear reaction term are present (multiplied for proper constants). In this subsection we write a generalization to $N\in \mathbb{N}$ species with arbitrary maximal occupation per site, again called $\nu\in\mathbb{N}_{0}$. The generator reads as follows

@@

    \mathcal{L}_{x,u}^{i}f(\bm{n})=\sum_{k,l=1}^{N}n_{k}^{x}n_{l}^{y}\left(f(\bm{n}-\delta_{k}^{x}+\delta_{g_{i}(l)}^{x}+\delta_{g_{i}(k)}^{y}-\delta_{l}^{y})-f(\bm{n})\right)

@@

    \mathcal{L}_{x,y}^{m}f(\bm{n})=\sum_{l,k=0}^{N}n_{k}^{x}\left(f(\bm{n}-\delta_{k}^{x}+\delta_{l}^{x})-f(\bm{n})\right)

@@

where with the symbol $\Delta_{N}$ we denote the discrete Laplacian operator on the chain, i.e. $\Delta_{L}n_{k}^{x}=n_{k}^{x+1}+n_{k}^{x-1}-2n_{k}^{x}$, $\forall x\in \{2,\ldots,L-1\}$ and $\forall k\in \{2,\ldots,N\}$. In the spirit of \cite{casini2022uphill} we can write he the Master's equation for the evolution of the average density occupation, obtaining a system of $N-1$ difference differential equation where the discrete Laplacian of all species are coupled and linear reaction term are present. \\ 

@@

where the duality matrix $D$ is again \eqref{dualityMatrix}. Following subsection \ref{subsectionDualityProof}, to prove \eqref{DualityRelationRD} we only need tho show that the operator $S_{x,y}$ defined in \eqref{symmetry} commutes with $\sum_{i=1}^{N}H_{x,y}^{i}$ and with $H_{x,y}^{m}$. This result will follow from the commutator between $\sum_{i=1}^{N}H_{x,y}^{i}$ and $\sum_{a=2}^{N}(E_{a1}^{x})\otimes \sum_{b=2}^{N}E_{b,1}^{x}$ and from the commutator between  and $\sum_{a=2}^{N}(E_{a1}^{x})\otimes \sum_{b=2}^{N}E_{b,1}^{x}$. Using \eqref{eq:comgl}, the bilinearity and associativity of the Kronecker product and the bilinearity of the brackets we obtain 

@@

Therefore, arguing similarly to subsection \ref{subsectionDualityProof}, \eqref{DualityRelationRD} can be proved. 
